---
title: 'Reproducible Research Peer Assessment #1'
author: "Tony Truong"
date: "September 18, 2015"
output: html_document
keep_md: true
---

#Peer Assignment # 1

#### Loading and Preprocessing the data

Data contained in file named activity.csv stored locally on my desktop

```{r}
setwd("C:/Users/Tony Truong/Desktop/Coursera R/Reproducable Research")
library(ggplot2)
ActData<-read.csv(file = "activity.csv", header = TRUE, sep = ",")
head(ActData)
```

Convert date from factor to date
```{r}
ActData$date<-as.Date(ActData$date, "%Y-%m-%d")
```



#### Create cDate and cDateInt columns to store POSIXct format date data (just in case)

```{r}
ActData$cDate<-as.POSIXct(ActData$date)
ActData$cDateInt<-(ActData$cDate+ActData$interval*60)
```

# What is the mean total number of steps taken per day?

Run crosstab to get total steps, with NA's omitted

```{r}
DailySteps<-xtabs(steps~cDate, data = ActData, exclude = NA)
```

Plot Total Daily Steps
```{r}
plot(DailySteps, type ="h", xlab="Date",main="Total Daily Steps, NAs excluded")
```

Calculate Mean & Median Daily Steps

```{r}
DailySteps<-data.frame(DailySteps)
MeanDailySteps<-mean(DailySteps$Freq)
MeanDailySteps
MedianDailySteps<-median(DailySteps$Freq)
MedianDailySteps
```
# What is the average total number of steps taken per day

Run crosstab with aggregate to create xtab with means (of intervals over all dates) as summary

```{r}
AvgDSteps<-xtabs(steps~interval, aggregate(steps~interval,ActData,mean, exclue =NA ))
plot(AvgDSteps, type = "l", ylab="Total Steps ", main="Total Steps per Interval")

```



#####Now we look for the interval that has the maximum average daily steps


```{r}
AvgDSteps<-data.frame(AvgDSteps)

MaxInterval<-AvgDSteps[which.max(AvgDSteps[,2]),]
MaxInterval
```
So the maximum average daily total steps occurs at interval 835 with a total average of 206.16 steps.


# Imputing missing value

#### Calculate the number of NA's in ActData
```{r}
colSums(is.na(ActData))
```
 I chose to substitute the interval average value for the NA value.
 So for each NA value found, must find the corresponding interval to lookup 
 the interval's daily average for substituting.

 Will create a new df so as not to overwrite original data.

```{r}
ActDataMod<- ActData
```
 Outer loop goes through all the rows in ActData looking for NA's
 Inner loop test for NA and lookup interval value in ActData 
 then the corresponding average interval steps is used to substitute for the NA 
 (also need to convert steps from factor back to integer) 
 print out modified data frame just to check....


```{r}
AvgDSteps<-data.frame(AvgDSteps)
for (i in 1:nrow(ActData)){

        if(is.na(ActData$steps[i])){
                ActDataMod$steps[i] <- subset(AvgDSteps, interval == ActData[i,3],select = Freq)
        }
}
ActDataMod$steps<-as.integer(ActDataMod$steps)
head(ActDataMod)
head(ActData)
```

#### Recalculate daily total steps with modified data
```{r}
DailyStepsMod<-xtabs(steps~cDate, data = ActDataMod, exclude = NA)
plot(DailyStepsMod, type ="h", xlab= "Date", ylab="Total Daily Steps", main= "NA's replaced with Interval Avg")

MeanDStepsMod<-mean(DailyStepsMod)
MeanDStepsMod
MedianDStepsMod<-median(DailyStepsMod)
MedianDStepsMod
```
The impact of missing values (NA's) in this case does NOT seem to be very significant. 

# Are there differences in activity patterns between weekdays and weekends?

### Adding column for Weekdays vs Weekend Activities
```{r}
WDaysList <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
ActDataMod$WDays<-factor((weekdays(ActDataMod$cDate) %in% WDaysList), levels=c(FALSE, TRUE), labels=c('weekend','weekday'))

```

# Weekday vs Weekend Activities

```{r}
WeekdaysActAvg<-data.frame(xtabs(steps~interval + WDays, aggregate(steps~interval + WDays, ActDataMod, mean)))

WeekdaysActAvg$interval<-as.integer(WeekdaysActAvg$interval)

g<-ggplot(WeekdaysActAvg, aes(x=interval, y=Freq))
g + facet_grid(WDays~.)+ geom_line()
```

It is quite clear from the plots that the activity levels for weekdays are quite distinct from weekend activities.   




